Toronto's Most Studious Wards: An Analysis of Annual Toronto Public Library's Circulation and Visits by Wards Pre and Post-COVID

### 2024-01-26 outputs/paper/paper.qmd
Final paper upload

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-circulation2018
#| fig-cap: Total 2018 Annual Circulation of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.2

@@

circulation_2018 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 1600000), breaks = seq(0, 1600000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Circulation")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-circulation2019
#| fig-cap: Total 2019 Annual Circulation of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.2

@@

circulation_2019 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
    scale_y_continuous(limits = c(0, 1600000), breaks = seq(0, 1600000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Circulation")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-circulation2020
#| fig-cap: Total 2020 Annual Circulation of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.2

@@

circulation_2020 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 1600000), breaks = seq(0, 1600000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Circulation")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-circulation2021
#| fig-cap: Total 2021 Annual Circulation of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.2

@@

circulation_2021 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 1600000), breaks = seq(0, 1600000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Circulation")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-circulation2022
#| fig-cap: Total 2022 Annual Circulation of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.2

@@

circulation_2022 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 2400000), breaks = seq(0, 2400000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Circulation")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-visits2018
#| fig-cap: Total 2018 Annual Visits of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.2

@@

visits_2018 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 2400000), breaks = seq(0, 2400000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Visits")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-visits2019
#| fig-cap: Total 2019 Annual Visits of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.2

@@

visits_2019 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 2400000), breaks = seq(0, 2400000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Visits")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-visits2020
#| fig-cap: Total 2020 Annual Visits of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.2

@@

visits_2020 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 2400000), breaks = seq(0, 2400000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Visits")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-visits2021
#| fig-cap: Total 2021 Annual Visits of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.2

@@

visits_2021 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 2400000), breaks = seq(0, 2400000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Visits")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-visits2022
#| fig-cap: Total 2022 Annual Visits of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.2

@@

visits_2022 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 2400000), breaks = seq(0, 2400000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Visits")

@@

### City-Wide Annual Circulation and Visits
Tables that provide a summary of the city-wide total visits (@tbl-tplannualvisits) and total circulation (@tbl-tplannualcirculation) for each year. Annual circulation and visitation metrics have high standard deviations of 5,502,796 and 6,349,144, respectively. The high standard deviations imply that the overall annual variance of these metrics is high, therefore indicating values are spread out from the mean. This is seen in the graphs for both usage metrics as 2020 and 2021 reveal significantly smaller values compared to 2018 and 2019. By viewing the annual data, it is clear that the graphs represent the annual data after comparing the table to the aforementioned sections.

@@

Although this report does not specifically focus on proving causation between the COVID-19 pandemic and library usage metrics, it is clear that the annual totals of these metrics do have differences pre and post-pandemic. The COVID-19 pandemic’s lockdown and social distancing measures could be a contributing factor to the sharp decline in total circulation and visits to libraries across all wards in Toronto seen in 2020 and 2021. As indicated by the Government of Ontario, public libraries were subject to lockdown measures which could have affected patrons' ability to visit and check out material. It could also be possible that patrons’ behaviour has changed post-2020 regarding public spaces and shared materials.

@@

>>>>

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-circulation2018
#| fig-cap: Total 2018 Annual Circulation of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.5

@@

# Create Bar Graph
circulation_2018 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 1600000), breaks = seq(0, 1600000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Circulation")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-circulation2019
#| fig-cap: Total 2019 Annual Circulation of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.5

@@

# Create Bar Graph
circulation_2019 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
    scale_y_continuous(limits = c(0, 1600000), breaks = seq(0, 1600000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Circulation")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-circulation2020
#| fig-cap: Total 2020 Annual Circulation of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.5

@@

# Create Bar Graph
circulation_2020 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 1600000), breaks = seq(0, 1600000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Circulation")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-circulation2021
#| fig-cap: Total 2021 Annual Circulation of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.5

@@

# Create Bar Graph
circulation_2021 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 1600000), breaks = seq(0, 1600000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Circulation")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-circulation2022
#| fig-cap: Total 2022 Annual Circulation of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.5

@@

# Create Bar Graph
circulation_2022 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 2400000), breaks = seq(0, 2400000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Circulation")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-visits2018
#| fig-cap: Total 2018 Annual Visits of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.5

@@

# Create Bar Graph
visits_2018 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 2400000), breaks = seq(0, 2400000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Visits")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-visits2019
#| fig-cap: Total 2019 Annual Visits of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.5

@@

# Create Bar Graph
visits_2019 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 2400000), breaks = seq(0, 2400000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Visits")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-visits2020
#| fig-cap: Total 2020 Annual Visits of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.5

@@

# Create Bar Graph
visits_2020 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 2400000), breaks = seq(0, 2400000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Visits")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-visits2021
#| fig-cap: Total 2021 Annual Visits of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.5

@@

# Create Bar Graph
visits_2021 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 2400000), breaks = seq(0, 2400000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Visits")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-visits2022
#| fig-cap: Total 2022 Annual Visits of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Ward
#| fig-pos: "H"
#| fig-height: 2.5

@@

# Create Bar Graph
visits_2022 |>
  ggplot(aes(x = ward_no, y = ward_total)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:25) +
  scale_y_continuous(limits = c(0, 2400000), breaks = seq(0, 2400000, by = 200000)) +
  labs(x = "Ward Number", y = "Total Annual Visits")

@@

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-annualcirculation
#| fig-cap: City-Wide Annual Circulation of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Year
#| fig-pos: "H"
#| fig-height: 2.5

# Create a new column of the annual_total values in millions
annual_circulation_data <- mutate(annual_circulation_data, million = annual_total/1000000)

annual_circulation_data |>
  ggplot(aes(x = year, y = million)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 2018:2022) + 
  scale_y_continuous(breaks = seq(1, 23, by = 1)) +
  labs(x = "Year", y = "Total Annual Circulation (In Millions)") +
  # Code referenced from: https://ggplot2.tidyverse.org/reference/geom_smooth.html
  geom_smooth(method = "loess", se = FALSE, color = "red")
```

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-annualvisits
#| fig-cap: City-Wide Annual Visits of Toronto Public Library Branches (Physical and Fixed Location Branches Only) by Year
#| fig-pos: "H"
#| fig-height: 2.5

# Create a new column of the annual_total values in millions
annual_visits_data <- mutate(annual_visits_data, million = annual_total/1000000)

annual_visits_data |>
  ggplot(aes(x = year, y = million)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_x_continuous(breaks = 2018:2022) + 
  scale_y_continuous(breaks = seq(1, 23, by = 1)) +
  labs(x = "Year", y = "Total Annual Visits (In Millions)") +
  # Code referenced from: https://ggplot2.tidyverse.org/reference/geom_smooth.html
  geom_smooth(method = "loess", se = FALSE, color = "red")
```

### City-Wide Annual Circulation and Visits
The graphs above provide a summary of the city-wide total visits (@tbl-tplannualvisits) and total circulation (@tbl-tplannualcirculation) for each year with the addition of a trend line. Annual circulation and visitation metrics have high standard deviations of 5,502,796 and 6,349,144, respectively. The high standard deviations imply that the overall annual variance of these metrics is high, therefore indicating values are spread out from the mean. This is seen in the graphs for both usage metrics as 2020 and 2021 reveal significantly smaller values compared to 2018 and 2019.

In @fig-annualcirculation, the trend line appears to highlight 2018 and 2019 having relatively annual circulation at roughly the same total followed by a significant decrease in 2020 with some recovery in latter years. This further corroborates the patterns seen in the annual circulation graphs by ward which also highlight this trough pattern. @fig-annualvisits shows a similar pattern with the exception of 2021 having the lowest total value. The pre and post-COVID-19 years for annual visits by year closely corresponed to the patterns seen in annual circulation by year. This is due to 2018-2019 being peaks with some increase in totals year-over-year after 2020 but not yet returned to pre-COVID-19 levels. By viewing the annual data, it is clear that the graphs representing the data by ward closely follow the patterns found in the data by year.

@@

Although this report does not specifically focus on proving causation between the COVID-19 pandemic and library usage metrics, it is clear that the annual totals of these metrics do have differences pre and post-pandemic. The COVID-19 pandemic’s lockdown and social distancing measures could be a contributing factor to the sharp decline in total circulation and visits to libraries across all wards in Toronto seen in 2020 and 2021. As indicated by the Government of Ontario, public libraries were subject to lockdown measures which could have affected patrons' ability to visit and check out material [@ongov]. It could also be possible that patrons’ behaviour has changed post-2020 regarding public spaces and shared materials.

@@

Reflecting on the secondary data that was conducted prior to the creation of this report, the data analysed does reflect the research findings of Jamali and Haider. They concluded that physical loans had significantly decreased [@libraryusage]. Although this report focused on all circulation associated to a Toronto Public Library branch with a physical and fixed location, it also found that circulation dropped by over 1,200,000 million in 2020 at its lowest in comparison to 2018. When considering the decrease in circulation and visits, this results in questions about overall quality of life Torontonians have post-COVID-19. As researched by Chow and Tian, public libraries are thought to improve different measures of quality of life [@libraryimpact]. With Torontonians visiting and borrowing from libraries in fewer numbers, it calls into question whether or not these usage patterns have also resulted in poorer social community health. Further research could be done to identify if there is a correlation between quality of life and library usage with special attention towards the social effects the pandemic due to the closure of libraries and other community services.
